---
title: "p8105_hw3_wz2506"
author: "Weiran Zhang"
date: "10/11/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(viridis)
library(forcats)
```


##Problem 1

```{r}
library(p8105.datasets)
data("instacart") ##Load data instacart

```

```{r}
##How many aisle in total
max(pull(instacart, aisle_id))
min(pull(instacart, aisle_id))

##wrangle the table to aisle_id and number of items ordered from each aisle
instacart = instacart %>%
  group_by(aisle_id) %>%
  summarize(n = n())


```
`Range of aisle_id is from 1 to 134, so there are 134 aisles in total.`


```{r}
##dataset of aisle_id and number of items ordered from the aisles over 10000 to check for the aisle with the most items ordered from
instacart %>%
  mutate(
    large = case_when(
      n > 10000 ~ "large",
      n <= 10000 ~ "not_large",
      TRUE     ~ ""
    )) %>%
  filter(large == "large") %>%
  group_by(aisle_id, large)

##plot of number of items ordered from each aisle for items number over 10000
instacart %>%
  mutate(
    large = case_when(
      n > 10000 ~ "large",
      n <= 10000 ~ "not_large",
      TRUE     ~ ""
    )) %>%
  filter(large == "large") %>%
  group_by(aisle_id, large) %>%
  ggplot(aes(x = aisle_id, y = n)) +
  geom_point()

```

`From the dataset of aisle_id and number of items over 10000, we can see that the aisle with aisle_id = 83 is the aisle with most items oredered from.`

```{r}
##Table of three most popular items in the three specific aisles
library(p8105.datasets)
data("instacart")
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarize(n = n()) %>%
  filter(min_rank(desc(n)) < 4 )%>%
  knitr::kable(digits = 1)

```

`So the three most popular items in aisle of baking ingredient are cane sugar(336), light brown sugar(499) and pure baking soda(387), in aisle of dog food care are organix chicken & brown rice recipe(28), small dog biscuits and snack sticks chicken(26) & rice recipe dog(30), in aisle packaged vegetables fruits are organic baby spinach(9784), organic blueberries(4966) and organic raspberries(5546). `

```{r}
##Table of mean hour of day
library(p8105.datasets)
data("instacart")
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  mutate(order_dow = replace(order_dow, order_dow == '0', 'Sunday'),
         order_dow = replace(order_dow, order_dow == '1', 'Monday'),
         order_dow = replace(order_dow, order_dow == '2', 'Tuesday'),
         order_dow = replace(order_dow, order_dow == '3', 'Wednesday'),
         order_dow = replace(order_dow, order_dow == '4', 'Thursday'),
         order_dow = replace(order_dow, order_dow == '5', 'Friday'),
         order_dow = replace(order_dow, order_dow == '6', 'Saturnday'),
  ) %>%
  group_by(product_name, order_dow) %>%
  summarize(meanhour = mean(order_hour_of_day)) %>%
  pivot_wider(
    id_cols = c(product_name, order_dow, meanhour),
    names_from = order_dow,
    values_from = meanhour) %>%
  knitr::kable(digits = 1)

```

`So from the table of mean hour of order, we can see that Wednesday it tooks longer to order for both Coffee ice Cream and Pink Lady Apples. But the mean hours are all about between 11-15 hours. `

`There are 15 variables in this instacart dataset and 1384617 observations in total. Key variables including aisle, aisle_id, product_name, order_dow, order_hour_of_hour.`

##Problem 2

```{r}
##Load BRFSS dataset
library(p8105.datasets)
data("brfss_smart2010")

```

```{r}
##Data cleaning following the instructions
tidy_data = brfss_smart2010 %>%
  filter(Topic == "Overall Health", 
         Response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(Response = as.factor(Response)) %>%
  mutate(Response = forcats::fct_relevel(Response, c("Excellent","Very Good","Good","Fair","Poor"))) %>%
  arrange(Response)
```

```{r}
##states that observed at 7 or more locations in year 2002
brfss_smart2010 %>%
  filter(Topic == "Overall Health", 
         Response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"), 
         Year == "2002") %>%
  group_by(Locationdesc) %>%
  separate(Locationdesc, c("state","county"), sep = "-") %>%
  group_by(state) %>%
  summarize(n = n()) %>%
  mutate(
    observed = case_when(
      n >= 7 ~ "observed",
      n < 7 ~ "not_observed",
      TRUE     ~ ""
    )) %>%
  filter(observed == "observed") %>%
  group_by(observed)

##states that observed at 7 or more locations in year 2010
brfss_smart2010 %>%
  filter(Topic == "Overall Health", 
         Response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"), 
         Year == "2010") %>%
  group_by(Locationdesc) %>%
  separate(Locationdesc, c("state","county"), sep = "-") %>%
  group_by(state) %>%
  summarize(n = n()) %>%
  mutate(
    observed = case_when(
      n >= 7 ~ "observed",
      n < 7 ~ "not_observed",
      TRUE     ~ ""
    )) %>%
  filter(observed == "observed") %>%
  group_by(observed)

```

```{r}
##“spaghetti” plot
brfss_smart2010 %>%
  select(Year, Response, Locationabbr, Data_value) %>%
  filter(Response == "Excellent") %>%
  group_by(Locationabbr, Year) %>%
  summarize(mean_data = mean(Data_value)) %>%
  ggplot(aes(x = Year, y = mean_data, color = Locationabbr)) +
  geom_point() + geom_line() 

```

```{r}
##Two-panel plots showing distribution of data_value
tidy_data %>%
  filter(Year %in% c(2006, 2010), 
         Locationabbr == "NY") %>%
  ggplot(aes(x = Response, y = Data_value)) +
  geom_boxplot() + 
  facet_grid(~Year) +
  viridis::scale_fill_viridis(discrete = TRUE)

```

##Problem 3

```{r}
##Load and tidy accel dataset
accel = read_csv(file = "/Users/weiranzhang/p8105_hw3_wz2506/accel_data.csv") %>%
  mutate(weekday_vs_weekend =
           ifelse(day %in% c("Saturnday","Sunday"), "Weekend", "Weekday"))
  
```

```{r}
##Total activity variable aggregate across minute
sum = accel %>%
  mutate(Total = rowSums(accel[,4:1443])) %>%
  select(week, day_id, day, Total)

```

```{r}
##single-panel plot of 24-hour activity vs. time
databyhour = accel %>%
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity"
  ) %>%
  mutate(minute = as.numeric(minute),
         hour = minute %/% 60) %>%
  group_by(day, hour, week) %>%
  summarize(activitycounts = sum(activity)) 

week1 = databyhour %>%
  filter(week == 1)

week2 = databyhour %>%
  filter(week == 2)

week3 = databyhour %>%
  filter(week == 3)

week4 = databyhour %>%
  filter(week == 4)

week5 = databyhour %>%
  filter(week == 5)

  ggplot() +
  geom_line(week1, mapping = aes(x = hour, y = activitycounts, color = day)) + geom_line(week2, mapping = aes(x = hour, y = activitycounts, color = day)) + geom_line(week3, mapping = aes(x = hour, y = activitycounts, color = day)) + geom_line(week4, mapping = aes(x = hour, y = activitycounts, color = day)) + geom_line(week5, mapping = aes(x = hour, y = activitycounts, color = day)) 

```




